---
title: "RetailCaseStudy"
output: html_document
---

#####  Library Used 
```{r message=FALSE,warning=FALSE} 
library(tidyverse)
library(lubridate)
library(knitr)
library(kableExtra)
```


#####  Loading the data  

```{r message=FALSE}

customer <- read_csv("pro data/Customer.csv")
product_hierarchy <- read_csv("pro data/prod_cat_info.csv")
transaction <- read_csv("pro data/Transactions.csv")

```

##### Q1 : Merge all the dataset,keeping all the customers who have done transaction

```{r }
# Part A ::Merging use base merge() function 

# Joining  customer and transaction table 

# Primary key : customer Id in Customer table 
# Foreign key : customer Id in Transactions table 

customer_trans <- merge(x = transaction,y = customer,
                    by.x = "cust_id",by.y = "customer_Id",all.x = T)

# Left join is used to include customer  transaction information

customer_final <-  merge(x = customer_trans,y = product_hierarchy,
                    by.x = c("prod_cat_code","prod_subcat_code"),
                    by.y = c("prod_cat_code","prod_sub_cat_code"),all.x = T) 

customer_final <- customer_final %>%   ## Removing the redundant data ##
                  select(cust_id,transaction_id,everything(),-prod_cat_code,-prod_subcat_code)    


knitr::kable(head(customer_final,10),caption = "Using Base merge") %>% kable_styling(full_width = T) 
# Part B :: Merging use dplyr merge function 

customer_final_dplyr <- left_join(x = transaction,y = customer,
                                  by = c("cust_id"="customer_Id")) %>%
                         left_join( product_hierarchy, 
                                  by = c("prod_cat_code" = "prod_cat_code",
                                  "prod_subcat_code" = "prod_sub_cat_code"))

customer_final_dplyr <- customer_final_dplyr %>%
                        select(everything(),-prod_subcat_code,-prod_cat_code)

knitr::kable(head(customer_final_dplyr,10),caption = "Using dplyr merge") %>% kable_styling(full_width = T) 
```

##### Ques 2: Prepare a summary report for the merged data

```{r}

# a: Column names and corresponding data types
dtypes <- data.frame(DataType = sapply(customer_final_dplyr,FUN = class)) 

knitr::kable(dtypes) %>% kable_styling(full_width = F,position = "left")

# b: Top/bottom 10 observations #

# Top 10
 top_10 <- customer_final_dplyr %>% head(10) 
 knitr::kable(top_10,caption = " Top 10 Observations ")%>% kable_styling(full_width = T) %>% kableExtra::scroll_box(height = "500px",width ="100%")

# Bottom 10  
bottom_10 <- customer_final_dplyr %>% tail(10) 
knitr::kable(bottom_10,caption = " Bottom 10 Observations ")%>% kable_styling(full_width = T) %>% kableExtra::scroll_box(height = "500px",width ="100%")


# Data manipulation 

# Data type conversion 
customer_final_dplyr$transaction_id <- as.character(customer_final_dplyr$transaction_id)
customer_final_dplyr$cust_id <- as.character(customer_final_dplyr$cust_id) 
customer_final_dplyr$city_code <- as.character(customer_final_dplyr$city_code) 

# Dividing data into Numerical and categorical

# Numeric
numeric_data <- sapply(X=customer_final_dplyr,FUN = is.numeric)
continuous_data <- customer_final_dplyr[,numeric_data]

# Categorical
character_data <- sapply(X=customer_final_dplyr,FUN = is.character)
freq_data <- customer_final_dplyr[,character_data]

# Handling Dates ##
customer_final_dplyr$tran_date <- lubridate::dmy(customer_final_dplyr$tran_date)
customer_final_dplyr$DOB <- lubridate::dmy(customer_final_dplyr$DOB)



# c: Five number summary for continuous variables
five_num_summary <- summary(continuous_data)

knitr::kable(five_num_summary,caption = " Five number summary ") %>% kable_styling(full_width = T)

# d: Frequency table for  categorical variable

list <- lapply(X=freq_data,FUN = plyr::count )

freq_table_gender <- list[["Gender"]] %>% rename("Gender"="x") 
knitr::kable(freq_table_gender,caption="Gender")%>% kable_styling(full_width = F,position = "left")

freq_table_prodCat <- list[["prod_cat"]] %>% rename("prod_cat"="x")
knitr::kable(freq_table_prodCat,caption="prod_cat")%>% kable_styling(full_width = F,position = "left")

freq_table_prodSubcat <- list[["prod_subcat"]] %>% rename("prod_subcat"="x")
knitr::kable(freq_table_prodSubcat,caption="prod_subcat")%>% kable_styling(full_width = F,position = "left")

freq_table_storeType <- list[["Store_type"]] %>% rename("Store_type"="x")
knitr::kable(freq_table_storeType,caption="Store_type")%>% kable_styling(full_width = F,position = "left")

freq_table_cityCode <- list[["city_code"]] %>% rename("city_code"="x")
knitr::kable(freq_table_cityCode,caption="city_code")%>% kable_styling(full_width = F,position = "left")%>% kable_styling(full_width = F,position = "left")

freq_table_CustID <-  list[["cust_id"]] %>% rename("CustId"="x") %>% arrange(desc(freq))
# Showing only top 10 rows
knitr::kable(head(freq_table_CustID,10),caption = "cust_id")%>% kable_styling(full_width = F,position = "left")


# NOTE: No frequency table for TransId, as every TransId is unique.


```
#### Ques 3A : Histograms for all continuous vars
```{r  message=F,fig.height=6.5,fig.align='center'}

cont_vars <- function(cont){
                    ggplot() +
                    geom_histogram(aes_string( x=cont),
                                   bins = 5,color="red",fill="lightblue")+
                    theme_bw()
                      }


cont_data_list <- lapply(continuous_data,FUN = cont_vars )

plot_Qty <- cont_data_list[["Qty"]] + labs(x="Quantity",title = "Qty Histogram")
plot_Rate <- cont_data_list[["Rate"]]+labs(x="Rate",title = "Rate Histogram")
plot_Tax <- cont_data_list[["Tax"]]+labs(x="Tax",title = "Tax Histogram")
plot_Total_amt <- cont_data_list[["total_amt"]]+labs(x="Total Amount",title = "Total_Amt Histogram")
  
cowplot::plot_grid(plot_Qty,plot_Rate,plot_Tax,plot_Total_amt)

```

#### Ques 3B : Frequency bars for all categorical vars.
```{r message=F,fig.height=6.5,fig.align='center'}
# Prod_cat
freq_data %>% ggplot(aes(x=prod_cat)) + geom_bar(fill = "steelblue",color="tomato",alpha=0.5)

# Gender
freq_data %>% ggplot(aes(x=Gender)) + geom_bar(fill = "steelblue",color="tomato",alpha=0.5)

#Store_type   
freq_data %>% ggplot(aes(x=Store_type)) + geom_bar(fill = "steelblue",color="tomato",alpha=0.5)

# City_code  
freq_data %>% ggplot(aes(x=city_code)) + geom_bar(fill = "steelblue",color="tomato",alpha=0.5)

# Prod_Subcat    
freq_data %>% ggplot(aes(y=prod_subcat)) + geom_bar(fill = "steelblue",color="tomato",alpha=0.5)
    
      
   

```


#### Ques4 :	Calculate the following information using the merged dataset :

```{r}
# A:	Time period of the available transaction data.
Time_period <- max(customer_final_dplyr$tran_date) - min(customer_final_dplyr$tran_date)
knitr::kable(Time_period,col.names = "No Of days")





# B:	Count of transactions where the total amount of transaction was negative.
negative_trans_count <-  sum(customer_final_dplyr$Qty<0,na.rm = T)
knitr::kable(negative_trans_count,col.names = "No Of Negative transactions")


```




#### Ques 5:	Analyze which product categories are more popular among females vs male customers.
    
```{r }
Male_Vs_Female <- customer_final_dplyr  %>% 
                  count(prod_cat,Gender,sort = T,name = "Count") %>% remove_missing(na.rm = T)


knitr::kable(Male_Vs_Female,caption ="Popular Product categories") %>% kable_styling(full_width = F,position = "left")


comp_graph <-  Male_Vs_Female %>%
                ggplot(aes(x=prod_cat,y=Count,fill=Gender))+
                geom_col(position = position_dodge())+
                labs(title = " Popular Product Category ",subtitle = " MALE VS FEMALE ")+
                theme(plot.title = element_text(size = 20,hjust = 0.5),
                      plot.subtitle = element_text(size=15,hjust = 0.5,face="bold"))

comp_graph

# Inference:
# Books, Electronics and Home and Kitchen,all are equally popular among male and female
# Footwear category is more popular among female as compared to male.
# Clothing is more popular among male as compared to female.

```


#### Ques 6 : 	Which City code has the maximum customers and the percentage of  customers from that city ?

```{r}

max_cust_city <-  customer_final_dplyr %>% 
                   filter(!is.na(city_code)) %>%
                   count(city_code,sort=T) %>%
                   mutate(prop = paste0 ( round(n/sum(n)*100,1),"%")) %>% 
                   head(1) 

knitr::kable(max_cust_city) %>% kable_styling(full_width = F,position = "left")

```


#### Ques 7 : 	Which store type sells the maximum products by value and by quantity?

```{r}
# Max products by quantity
max_prd_by_qty <- customer_final_dplyr %>% count(Store_type,wt=Qty,sort = T) %>% head(1)

knitr::kable(max_prd_by_qty) %>% kable_styling(full_width = F,position = "left")

# Max products by amount
max_prd_by_amt <-  customer_final_dplyr %>% count(Store_type,wt=total_amt,sort = T) %>% head(1)

knitr::kable(max_prd_by_amt) %>% kable_styling(full_width = F,position = "left")

```


#### Ques 8: What was the total amount earned from the "Electronics" and "Clothing" categories from Flagship Stores?

```{r message=FALSE}

       
total_by_prod_cat <-  customer_final_dplyr %>% 
                      filter(
                      Store_type == "Flagship store",
                      prod_cat %in% c("Electronics","Clothing")) %>% 
                      group_by(Store_type,prod_cat) %>%
                      summarize(Total_sales = sum(total_amt))

knitr::kable(total_by_prod_cat) %>% kable_styling(full_width = F,position = "left")

```


#### Ques 9: What was the total amount earned from "Male" customers under the "Electronics" category ?

```{r}

total_amt_earned <- customer_final_dplyr %>%
                      filter(
                      Gender == "M",
                      prod_cat == "Electronics") %>%
                      group_by(prod_cat) %>% 
                      summarize(Total_Amt_Earned = prettyNum(sum(total_amt),
                      big.mark = ",",decimal.mark = "."))%>%
                      select(Total_Amt_Earned)


knitr::kable(total_amt_earned,col.names = "Total Amt Earned")

```



#### Ques 10: How many customers have more than 10 unique transactions, after removing all transactions which have any negative amounts?
 
```{r}
  
unique_10_trans <- customer_final_dplyr %>% 
                    filter(total_amt>0) %>%  
                    group_by(cust_id) %>% 
                    summarise(Count=n()) %>% 
                    filter(Count>=10) %>% 
                    arrange(desc(Count)) %>% 
                     nrow()

knitr::kable(unique_10_trans,col.names = "No Of Unique Transactions")


```


#### Ques 11:	For all customers aged between 25 - 35, find out:

```{r}

### Data To analyze
ques11 <- customer_final_dplyr %>%
          mutate(
          age = eeptools::age_calc(DOB,units = "years",precise = T)) %>% 
          filter(
          age >=25 & age<=35) %>% 
          mutate(Age=round(age)) %>%
          select(-age)

# a: What was the total amount spent for “Electronics” and “Books” product categories?

Ans11_A <- ques11 %>% 
            filter(
            prod_cat %in% c("Electronics","Books")) %>%
            group_by(prod_cat) %>% 
            summarize(Total=sum(total_amt))

knitr::kable(Ans11_A) %>% kable_styling(full_width = F,position = "left")

# b: What was the total amount spent by these customers between 1st Jan,2014 to 1st Mar,2014?
Ans11_B <-  ques11 %>%
              filter(
              tran_date >= "2014-01-01" & tran_date <= "2014-03-01") %>%
              summarise(Total_Amt_Spent=sum(total_amt))
Ans11_B
knitr::kable(Ans11_B) %>% kable_styling(full_width = F,position = "left")



```




